In the modern world, software is ubiquitous: from the personal computers most of
us use to the mission critical systems that require fine control. Software
quality is usually a result of a healthy software project, \ie it is not just a
function of the code, rather also of the process producing that code.
Further, projects are not just source-code; indeed, they contain a wealth of
documents written in English or other languages. Focusing strictly on
source-code tells us only half the story.

In this thesis, I propose the notion of project health; a notion that was
previously colloquially understood and that encompasses those processes that
help project succeed and flourish. Under this umbrella, I focus on three tasks
that could impede project success: pull-request-issue linking, commit separation
into atomic patches, and mixed-text segmentation and pre-processing. I focus on
tasks that would break assumptions made by researchers when proposing techniques
in the first two, while the latter presents a new way of handling data enabling
analysis which is aware of algorithmic and natural language information. By
borrowing techniques and methodology from Machine Learning, I propose prototype
systems to resolve these issues.

The work I present in this thesis follows the ethos of helping developers help
themselves and us. We, as researchers, help developers maintain project health
with its inherent benefits: easier onboarding, lower costs of maintenance \etc
This in turn can create better training data for researchers and may enable yet
better automation techniques and research avenues for Big Code.